---
title: "Documenting Monitor"
author: "Minoo"
date: "2022-11-30"
output: word_document
---

The Monitor data has been received on 18.08.2022 as zip file called Minoo, and is saved in this pathway C:\\Users\\PhysioUser\\Desktop\\PhD\\Monitor A copy of the raw signals has been saved in Monitor\\raw signals. then the signals are preprocessed and saved in 3 subfolders in Monitor\\pre-processed signals:

-   Filtered signals
-   Nirs containing transients
-   Transients

The code used to pre-process the raw NIRS signals is Monitor.m. It has been modified for each participant based on individual differences. the outcome is saved as mydata.txt.

Then transient_extraction.m has been used to extract the transients and saved files as transient and filtered.

## Transients
### Extracting features
the transient_feature_extraction.m file has been used to extract the features from the transients. The outcome is saved as transient_features_MN. The code transient_feature_extraction_below63.m is used to calculate the time rcSO2 signal is spending below 63%, and transient_features_extraction_below63_transient.m has been used to calculate the time transient signal is spending below 63%. calculating_nirs_duration.m has been used to have the duration of NIRS in each participant. my_code_put_together_allfeatures.m has been used to gather all the features extracted from transients in a single .csv file called all_transient_features.csv.

```{r all_feature, echo=FALSE, include=FALSE}
transient_all <- readxl::read_excel("all_transient_features.xlsx")

```

```{r table, echo=FALSE, include=FALSE}
knitr::kable(transient_all, digits =  4, 
             caption = "The features extracted from transients in all participants")
```

 
### Optimization for xgboost parameters_MRI
61 infants were recruited and NIRS was measured for them, when the NIRS had been pre-processed, because some files had to be separated to two, we had 65 files. MN0052 was excluded since he had the diagnosis of sepsis, and we could not say for definite that HIE was the cause of their encephalopathy. MN0030 and MN0041 were mild that had been cooled, so they do not have MRI outcome but could be included in the HIE group. For MRI outcome, we have 58 infants, and 62 files, since 4 babies had 2 files. We have averaged the values except for the time spent below 63%, which was added, and we have 58 rows in total. 

Then a leave-one-out method optimization has been done to find the best values for xgboost method. the python code used is optimization.IPYNB. The transient features are saved as *transient for xgb_MRI.csv* which is colab-friendly! The result is saved in optimization_transient_xgb.csv. The outcome used is *Composite outcome of barkovich MRI abnormality +/- death (0=normal, 1=abnormal)* which has been in the file called  MRI outcome for xgb_58ver.csv. The optimized values are:

```{r MRI optimization, echo=FALSE, include=FALSE}
MRI_optimization <- readxl::read_excel("optimazation_transient_MRI_xgb.xlsx")

```

```{r optimized parameters for MRI, echo=FALSE, include=TRUE}
knitr::kable(MRI_optimization[14,], digits =  4, 
             caption = "The optimized parameters for XGBoost with MRI outcome")
```

### leave one out xgboost_MRI
The code called transient_xgboost.IPYNB is used for the leave one out xgboost method.the files used for this purpose are MRI outcome for xgb_58ver.csv and transient for xgb_MRI.csv. Then the matlab code auc_CI_sen_spec.m has been used to calculate the 95% confidence interval of AUC, accuracy, sensitivity and specificity:  
```{r monitor probabilities and AUC, echo=FALSE, include=FALSE}
monitor_prob_auc <- readxl::read_excel("Monitor_probabilities_AUC.xlsx")
```

```{r tran auc, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,1:2], digits =  4, 
             caption = "The AUC for transients with MRI outcome")
```

To be sure about the used parameters, we take another set of them to see if there is any difference in AUC or not, so we have: 

```{r another optimized parameters for MRT, echo=FALSE, include=TRUE}
knitr::kable(MRI_optimization[26,], digits =  4, 
             caption = "The optimized parameters for XGBoost with MRI outcome")
```

Using different parameters, but the same code and files, still the AUC was not significant: 

```{r another tran auc, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,3)], digits =  4, 
             caption = "The AUC for transients with MRI outcome")
```

### MRI watershed outcome
We used the same parameters as MRI, the files transient for xgb_MRI.csv, watershed outcome for xgb.58ver.csv, and transient_xgboost.IPYNB, and the AUC was:

```{r, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,14)], digits =  4, 
             caption = "The AUC for transients with MRI_watershed outcome")
```

### MRI gray injury outcome
We used the same parameters as MRI, the files transient for xgb_MRI.csv, gray injury outcome for xgb.58ver.csv, and transient_xgboost.IPYNB, and the AUC was:

```{r, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,15)], digits =  4, 
             caption = "The AUC for transients with MRI_watershed outcome")
```

### Optimization for xgboost parameters_HIE
61 infants were recruited and NIRS was measured for them, when the NIRS had been pre-processed, because some files had to be separated to two, we had 65 files. MN0052 was excluded since he had the diagnosis of sepsis, and we could not say for definite that HIE was the cause of their encephalopathy. MN0030 and MN0041 were mild that had been cooled, and were included in the HIE group with non-injury label. For HIE outcome, we have 60 infants in total, some files are separated,so we have 64 files, but at the end the features are added and averaged for babies with more than one file and we have 60 rows at the end. 

Then a leave-one-out method optimization has been done to find the best values for xgboost method. the python code used is optimization.IPYNB. The transient features are saved as *transient for xgb_HIE.csv* which is colab-friendly! The result is saved in optimization_transient_HIE_xgb.csv. The outcome used is *Clinical Grade of HIE (0=mild, 1=moderate or severe)* which has been in the file called  HIE outcome for xgb_60ver.csv. The optimized values are:

```{r HIE optimization, echo=FALSE, include=FALSE}
HIE_optimization <- readxl::read_excel("optimization_transient_HIE_xgb.xlsx")

```

```{r optimized parameters for HIE, echo=FALSE, include=TRUE}
knitr::kable(HIE_optimization[3,], digits =  4, 
             caption = "The optimized parameters for XGBoost with HIE outcome")
```

### leave one out xgboost_HIE
The code called transient_xgboost.IPYNB is used for the leave one out xgboost method.the files used for this purpose are HIE outcome for xgb_60ver.csv and and transient for xgb_HIE.csv. the AUC is significant: 

```{r tran auc HIE, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,4)], digits =  4, 
             caption = "The AUC for transients with HIE outcome")
```

To be sure about the used parameters, we take another set of them to see if there is any difference in AUC or not, so we have: 

```{r another optimized parameters for HIE, echo=FALSE, include=TRUE}
knitr::kable(HIE_optimization[25,], digits =  4, 
             caption = "The optimized parameters for XGBoost with HIE outcome")
```

Using different parameters, but the same code and files, still the AUC was significant: 

```{r another tran auc HIE, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,5)], digits =  4, 
             caption = "The AUC for transients with HIE outcome")
```

In the next level, we were wondering if the features extracted from the transients could predict the HIE level, so we used the same files, just in the HIE outcome for xgb_60ver, instead of column 2, we used the HIE levels in column 3. for optimization and training the model, we used multiclass_xgboost.ipynb python file. The optimized parameters are: 

```{r, echo=FALSE, include=FALSE}
HIE_level_optimization <- readxl::read_excel("optimization_transient_HIE_grading.xlsx")

```

```{r, echo=FALSE, include=TRUE}
knitr::kable(HIE_level_optimization[3,], digits =  4, 
             caption = "The optimized parameters for XGBoost with HIE grading outcome")
```

Then, using the R code called "multiclass_auc&ci", and the probabilities saved in the file multiclass_prob.xlsx, we calculated the AUC and 95% confidence interval. 

```{r, echo=FALSE, include=FALSE}
multiclass <- readxl::read_excel("Monitor_probabilities_AUC.xlsx", sheet = "multiclass")

```

```{r, echo=FALSE, include=TRUE}
knitr::kable(multiclass[1:3, c(1,2)], digits =  4, 
             caption = "The AUC and 95% CI for transient with HIE grading outcome")
```

## NIRS signal containing tarnsients
### Extracting features
the nirs_feature_extraction.m file has been used to decompose the signal into 5 bandwidths, divide them into epochs of 4 hours with 2 hours of overlapping and extract the features from each epoch in each bandwidth of the NIRS signal that still contains transients. The outcome is saved as nirs_features_MN. The extracted features are: 

- mean envelope (*5)
- standard deviation envelope (*5)
- signal kurtosis (*5)
- signal skewness (*5)
- 5th percentile of envelope (*5)
- 95th percentile of envelope (*5)
- mean instantaneous frequency (*5)
- standard deviation instantaneous frequency (*5)
- kurtosis instantaneous frequency (*5)
- skewness instantaneous frequency (*5)
- 5th percentile instantaneous frequency (*5)
- 95th percentile instantaneous frequency (*5)
- fractal dimension (*5)
- the energy of signal in 5th percentile of the signal duration
- the energy of signal in 95th percentile of the signal duration

This gives us 67 features in total, but we will not include the energy of the signal in the 5th and 95th percentile of the signal duration. 

my_code_put_together_allfeatures_nirs.m has been used to gather all the features extracted from NIRS in a single .csv file called all_nirs_features.csv.

### Optimization for NIRS xgboost parameters_MRI
Same as what was explained in *optimization for xgboost parameters_MRI*, we had 62 files at the end.Since each NIRS recording has different duration, we had different number of epochs for each infant. the colab friendly file containing the info for these 62 files is saved as *nirs with MRI outcome for xgb*. it is a 1505*66 matrix. The first column contains the IDs and columns 2-66 correspond to the 65 features for each epoch. The outcome is saved as MRI outcome for xgb_nirs.csv.

Then a leave-one-out method optimization has been done to find the best values for xgboost method. the python code used is optimization.IPYNB. MRI outcome for xgb_58ver.csv has been used for getting AUC. The optimized values are:

```{r MRI optimization nirs, echo=FALSE, include=FALSE}
MRI_optimization_nirs <- readxl::read_excel("optimization_nirs_MRI_xgb.xlsx")

```

```{r optimized parameters for MRI nirs, echo=FALSE, include=TRUE}
knitr::kable(MRI_optimization_nirs[3,], digits =  4, 
             caption = "The optimized parameters for XGBoost with MRI outcome in nirs")
```

### leave one out xgboost_MRI_NIRS
The code called transient_xgboost.IPYNB is used for the leave one out xgboost method.the files used for this purpose are nirs with MRI outcome for xgb.csv, MRI outcome for xgb_nirs.csv and MRI outcome for xgb_58ver.csv (the last one used for getting AUC). The AUC is not significant: 

```{r nirs auc, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,6)], digits =  4, 
             caption = "The AUC for NIRS with MRI outcome")
```

To be sure about the used parameters, we take another set of them to see if there is any difference in AUC or not, so we have: 

```{r another optimized parameters for MRI NIRS, echo=FALSE, include=TRUE}
knitr::kable(MRI_optimization_nirs[10,], digits =  4, 
             caption = "The optimized parameters for XGBoost with MRI outcome in NIRS")
```

Using different parameters, but the same code and files, still the AUC was not significant: 

```{r another nirs auc, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,7)], digits =  4, 
             caption = "The AUC for NIRS with MRI outcome")
```

### Optimization for NIRS xgboost parameters_HIE
Same as what was explained in *optimization for xgboost parameters_HIE*, we had 64 files at the end.Since each NIRS recording has different duration, we had different number of epochs for each infant. the colab friendly file containing the info for these 64 files is saved as *nirs with HIE outcome for xgb*. it is a 1549*66 matrix. The first column contains the IDs and columns 2-66 correspond to the 65 features for each epoch. The outcome is saved as HIE outcome for xgb_nirs.csv.

Then a leave-one-out method optimization has been done to find the best values for xgboost method. the python code used is optimization.IPYNB. HIE outcome for xgb_60ver.csv has been used for getting AUC. The optimized values are:

```{r HIE optimization nirs, echo=FALSE, include=FALSE}
HIE_optimization_nirs <- readxl::read_excel("optimization_nirs_HIE_xgb.xlsx")

```

```{r optimized parameters for HIE nirs, echo=FALSE, include=TRUE}
knitr::kable(HIE_optimization_nirs[4,], digits =  4, 
             caption = "The optimized parameters for XGBoost with HIE outcome in nirs")
```

### leave one out xgboost_HIE_NIRS
The code called transient_xgboost.IPYNB is used for the leave one out xgboost method.the files used for this purpose are nirs with HIE outcome for xgb.csv, HIE outcome for xgb_nirs.csv and HIE outcome for xgb_60ver.csv (the last one used for getting AUC). The AUC is significant: 

```{r, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,8)], digits =  4, 
             caption = "The AUC for NIRS with HIE outcome")
```

To be sure about the used parameters, we take another set of them to see if there is any difference in AUC or not, so we have: 

```{r, echo=FALSE, include=TRUE}
knitr::kable(HIE_optimization_nirs[25,], digits =  4, 
             caption = "The optimized parameters for XGBoost with HIE outcome in NIRS")
```

Using different parameters, but the same code and files, still the AUC was significant: 

```{r, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,9)], digits =  4, 
             caption = "The AUC for NIRS with HIE outcome")
```

## Filtered NIRS signal
### Extracting features
the nirs_feature_extraction.m file has been used to decompose the signal into 5 bandwidths, divide them into epochs of 4 hours with 2 hours of overlapping and extract the features from each epoch in each bandwidth of the Filtered signal. The outcome is saved as filtered_features_MN.txt. The extracted features same as extracted features from the NIRS signals containing transients. The code my_code_put_together_allfeatures_nirs.m has been used to put all the features in a single file called all_filtered_features.csv. 

### Leave one out xgboost_MRI_filtered
The same parameters used for the NIRS file will be used for training these signals, so we have used these set of parameters: 

```{r, echo=FALSE, include=TRUE}
knitr::kable(MRI_optimization_nirs[3,], digits =  4, 
             caption = "The optimized parameters for XGBoost with MRI outcome in nirs")
```

and 

```{r, echo=FALSE, include=TRUE}
knitr::kable(MRI_optimization_nirs[10,], digits =  4, 
             caption = "The optimized parameters for XGBoost with MRI outcome in NIRS")
```

for the MRI outcome, and using the same code transient_xgboost.IPYNB, and the files called filtered with MRI outcome for xgb.csv, MRI outcome for xgb_nirs.csv and MRI outcome for xgb_58ver.csv have been used for training and testing the model, and this is the result: 

```{r, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,10)], digits =  4, 
             caption = "The AUC for Filtered with MRI outcome")
```

and 

```{r, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,11)], digits =  4, 
             caption = "The AUC for Filtered with MRI outcome")
```

### Leave one out xgboost_HIE_filtered
The same parameters used for the NIRS file will be used for training these signals, so we have used these set of parameters: 

```{r, echo=FALSE, include=TRUE}
knitr::kable(HIE_optimization_nirs[4,], digits =  4, 
             caption = "The optimized parameters for XGBoost with HIE outcome in nirs")
```

and 

```{r, echo=FALSE, include=TRUE}
knitr::kable(HIE_optimization_nirs[25,], digits =  4, 
             caption = "The optimized parameters for XGBoost with HIE outcome in NIRS")
```

for the HIE outcome, and using the same code transient_xgboost.IPYNB, and the files called filtered with HIE outcome for xgb.csv, HIE outcome for xgb_nirs.csv and HIE outcome for xgb_60ver.csv have been used for training and testing the model, and this is the result: 

```{r, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,12)], digits =  4, 
             caption = "The AUC for Filtered with HIE outcome")
```

and 

```{r, echo=FALSE, include=TRUE}
knitr::kable(monitor_prob_auc[61:65,c(1,13)], digits =  4, 
             caption = "The AUC for Filtered with HIE outcome")
```
